---
title: "ECHO Split-Half Reliablity Script"
author: "Santi"
date: "2/5/2019"
output:
  html_document:
    toc: true
    fig_height: 8.5
    fig_width: 12
    css: custom 2.css
---

This script computes the effect sizes and split-half reliablity for different ERP conditions in increasing numbers of trials. 

```{r, echo = FALSE}
options(width = 160)
options(knitr.table.format = "html")
```

# Setup
```{r eval=TRUE, echo=T, message=FALSE, warning=FALSE, results='hide', fig.show='hide'}
# Set working directory
setwd("/Users/santiagomorales/Dropbox/Fox_Lab/TOTS/CBCL_Trajectories/") 

list.of.packages <- c("psych", "zoo", "reshape2", "car","taRifx", "ggplot2", "nlme", "R.matlab", "tidyr","dplyr","foreach", "doParallel","effsize")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
# Loading packages
lapply(list.of.packages, require, character.only = TRUE)

# Setting my plotting options
my_opts <- list(theme_classic() + theme(axis.text=element_text(size=14), axis.title=element_text(size=15,face="bold"), legend.title=element_text(size=14,face="bold"), legend.text=element_text(size=13), strip.text.x = element_text(size = 14, face="bold")))
source("/Users/santiagomorales/Dropbox/R_dropbox/data.check_function.R") # Reading in function from file
############################################################

# Options for parallel
registerDoParallel(4)  # use multicore, set to the number of our cores
opts <- list(chunkSize=2)
comb <- function(x, ...) { # Setting up this function for later
  lapply(seq_along(x),
    function(i) c(x[[i]], lapply(list(...), function(y) y[[i]])))
}

```

The data should be trial-level data for each subject in one large spreadsheet. The script here assumes that the participant is labeled "id" and the different conditions as "Condition" and the different groups (in this case different ages) "age."

# Reliability
## ERPs
#### Between age
```{r}
dfm.ag <- dfm.ag %>% unite(Condition, Condition, age, remove = T)  # To look across age, I need to add age here to the condition
dfm.ag$Condition <- as.factor(dfm.ag$Condition) # This variable needs to be a factor

system.time({
rel_result_age <- foreach(s = seq(4, 32, 4), .combine='comb', .options.nws=opts, .multicombine=TRUE,  .errorhandling="pass",
                .init=list(list(), list())) %:%           # If adding a third list, remember to add a list here e.g., .init=list(list(), list(), list()))
  foreach(i = 1:3000, .combine='comb', .multicombine=TRUE,  .errorhandling="pass",  # Indicating how many subsamples
                .init=list(list(), list())) %dopar% {  # If adding a third list, remember to add a list here e.g., .init=list(list(), list(), list()))
    seed = sample(1:10000000, 1)
    set.seed(seed)   ## set the seed to make your partition reproducible
    
    list.of.packages <- c("psych", "zoo", "reshape2", "car","taRifx", "ggplot2", "tidyr","dplyr","foreach", "doParallel","effsize")       
    lapply(list.of.packages, require, character.only = TRUE)  
    
    # Before subsampling, I need to delete participants that do not have enough trials
    n_trials <- dfm.ag %>%
      group_by(Condition, id) %>%
      tally() 
    
    ids.out <- n_trials[n_trials$n < s,] %>% dplyr::select(Condition, id) %>% unite(Condition_id, Condition, id)

    # # This is similar to the loop above however, I need to create a subsampled dataset first
    df_temp_s <- dfm.ag %>%
      unite(Condition_id, Condition, id, remove = F) %>%
      dplyr::filter(!Condition_id %in% ids.out$Condition_id) %>% 
      group_by(Condition, id) %>% 
      sample_n(s)
    
    # Now I can do the split half on the subsampled dataset
    df_temp <- df_temp_s %>% 
      sample_frac(.50) %>%      # Creating a random half
      mutate(bin = 1) %>%       # Indicating that this is the first half 
      dplyr::select(id, trial,Condition, bin) %>%  # Keeping only vars of interest
      right_join(df_temp_s, by = c("id", "trial", "Condition")) %>%  # Bringing in the other half of the trials
      mutate(bin = if_else(is.na(bin), 2, 1)) %>%  # Creating the index for the second half of the trials
      group_by(Condition, id, bin) %>%    # Grouping by vars of interest
      dplyr::summarise(Mean_amp = mean(value, na.rm = T)) %>% # Getting the mean amplitude by vars of interest
      unite(Condition_bin, Condition, bin)  %>% # Creating the variable name
      spread(Condition_bin, Mean_amp) # Going to wide using our new variable
    
    
    # Checking the reliability
    r_temp <- corr.test(df_temp[,-c(1)])$r
    n_temp <- corr.test(df_temp[,-c(1)])$n
    r_temp <- (2*r_temp)/(1 + abs(r_temp)) # SB formula
    
    conditions_temp <- unique(gsub("(.*)_.*","\\1",row.names(r_temp))) # removing everything after the last "_" and only keeping unique conditions  
    
    r_temp_df <- data.frame()
    for (j in seq(2,sqrt(length(r_temp)), 2) ) { # Setting up a loop over the number of conditions
      r_temp_df_j <- r_temp[j,j-1]               # getting the index for that condition it is the row number and the column of j - 1
      r_temp_df_j <- data.frame(variable = conditions_temp[j/2], rcoeff = r_temp_df_j) # Getting the name of the conditions and it has to be divided by two
      r_temp_df_j$n <- ifelse(length(n_temp) == 1, n_temp, ifelse(length(n_temp) > 1, n_temp[j,j-1], "Error!")) # getting the n for that condition it is the same for all or the row number and the column of j - 1
      r_temp_df_j$seed = seed # Setting the seed 
      r_temp_df_j$n_trials = s # Setting the number that was subsampled by
      
      r_temp_df <- bind_rows(r_temp_df, r_temp_df_j)
    
    }
    
    ##### ##### ##### ##### ##### ##### ##### ##### ##### ##### ##### ##### #####
    # Saving the t-value and effect size to determine the power to detect significant Condition effect
    # Now using the subsampled dataset to get it ready for the t-test and effect size
    df_t <- data.frame()
    df_t_temp <- df_temp_s %>%
      group_by(Condition, id) %>%
      dplyr::summarise(Mean_amp = mean(value, na.rm = T)) # %>% # Next I just have to spread the data and run the t-test
      # spread(Condition, Mean_amp) # Going to wide using our new variable
    t_temp <- data.frame()
    for (age_i in c("4Y","5Y","7Y","9Y")) {
      df_t_temp_age <- df_t_temp[grepl(age_i, df_t_temp$Condition),]
      df_t_temp_age$Condition <- gsub(paste0("_",age_i), "", df_t_temp_age$Condition)
      # df_SME_temp_age <- df_t_temp_age # Saving this for SME computation
            
      df_t_temp_age <- spread(df_t_temp_age, Condition, Mean_amp)
      t_result_ERN <- t.test(df_t_temp_age$Correct_ERN, df_t_temp_age$Error_ERN, paired = T)
      eff_result_ERN <- cohen.d(df_t_temp_age$Correct_ERN, df_t_temp_age$Error_ERN, paired = T, na.rm=T)
      t_result_Pe <- t.test(df_t_temp_age$Error_Pe, df_t_temp_age$Correct_Pe, paired = T)
      eff_result_Pe <- cohen.d(df_t_temp_age$Error_Pe, df_t_temp_age$Correct_Pe, paired = T, na.rm=T)
      
      t_temp_age <- data.frame(t_val_ERN = t_result_ERN$statistic, df_ERN = t_result_ERN$parameter, Mean_dif_ERN = t_result_ERN$estimate,
                               Cohend_ERN = eff_result_ERN$estimate, Cohend.LL_ERN = eff_result_ERN$conf.int[1], Cohend.UL_ERN = eff_result_ERN$conf.int[2],
                         t_val_Pe = t_result_Pe$statistic, df_Pe = t_result_Pe$parameter, Mean_dif_Pe = t_result_Pe$estimate,
                               Cohend_Pe = eff_result_Pe$estimate, Cohend.LL_Pe = eff_result_Pe$conf.int[1], Cohend.UL_Pe = eff_result_Pe$conf.int[2],
                         seed = seed, n_trials = s, Age = age_i)
      
      t_temp <- bind_rows(t_temp, t_temp_age)
    }
    
    df_t <- bind_rows(df_t, t_temp)
    
    # ##### ##### ##### ##### ##### ##### ##### ##### ##### ##### ##### ##### #####
    # # Computing the SME based on Luck et al 2020 (preprint)
    # # Now using the subsampled dataset to get it ready for the SME computation for each participant
    # df_SME_temp <- df_temp_s %>%
    #   group_by(Condition, id) %>%
    #   dplyr::summarise(SME = (sd(value, na.rm = T)/sqrt(length(value)) ),
    #                    seed = seed, n_trials = s)  # %>% # Next I just have to spread the data and run the t-test
    
    # Returning everything out of the parallel loop
    # return(list(r_temp_df, df_t, df_SME_temp)) #If adding a third list, remember to add a list on loops above e.g., .init=list(list(), list(), list()))
    return(list(r_temp_df, df_t))
}
})
df_r <- do.call(dplyr::bind_rows, rel_result_age[[1]])
df_t <- do.call(dplyr::bind_rows, rel_result_age[[2]])
# df_SME <- do.call(dplyr::bind_rows, rel_result_age[3]) # Maybe this is not very efficient and maybe not useful

```

##### Plots
```{r}
cutoff <- 6 # Indicating how many participants should be there to believe the correlation

df_r$Age <- as.numeric(gsub("[^\\d]+", "", df_r$variable, perl=TRUE))# Separating age and condition
df_r$variable <- sub("_[^_]+$", "", df_r$variable)
# Making sure that they at least have 6 people before looking at their correlation
df_r$rcoeff[df_r$n < cutoff] <- NA
df_r$rcoeff[df_r$rcoeff < 0] <- 0


dfm_r <- df_r # no need for melting anymore

# Plotting
ggplot(dfm_r, aes(n_trials, rcoeff, color = variable, fill = variable, group = variable)) + stat_summary(fun.y = mean, geom = "point",  position=position_dodge(width=0.7)) + stat_summary(fun.y = mean, geom = "line",  aes(group = variable)) + my_opts+ labs(x = "# of Trials", y = "Split-Half Reliability")  + geom_hline(yintercept=.9, color='black', linetype = 2) + geom_hline(yintercept=.8, color='black', linetype = 1) + geom_hline(yintercept=.7, color='black', linetype = 3) + geom_hline(yintercept=.6, color='red', linetype = 4) +  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, by = 5)) + theme(legend.position="bottom", legend.title = element_blank()) + facet_wrap(~Age)

ggplot(dfm_r, aes(n_trials, rcoeff, color = as.factor(Age), fill = as.factor(Age), group = as.factor(Age))) + stat_summary(fun.y = mean, geom = "point",  position=position_dodge(width=0.7)) + stat_summary(fun.y = mean, geom = "line",  aes(group = Age)) + my_opts+ labs(x = "# of Trials", y = "Split-Half Reliability")  + geom_hline(yintercept=.9, color='black', linetype = 2) + geom_hline(yintercept=.8, color='black', linetype = 1) + geom_hline(yintercept=.7, color='black', linetype = 3) + geom_hline(yintercept=.6, color='red', linetype = 4) +  scale_x_continuous(limits = c(0, 34), breaks = seq(0, 32, by = 4)) + theme(legend.position="bottom", legend.title = element_blank()) + facet_wrap(~variable)


# Plots with confidence intervals
dfm_r_ag <- dfm_r %>% 
  group_by(n_trials, variable, Age) %>% # Grouping by vars of interest
  dplyr::summarise(Mean = mean(rcoeff, na.rm = T), Median = median(rcoeff), ci_L = quantile(rcoeff, .025, na.rm = T), ci_U = quantile(rcoeff, .975, na.rm = T))

# dfm_r_ag$ci_L[dfm_r_ag$variable != "Updated"] <- NA
# dfm_r_ag$ci_U[dfm_r_ag$variable != "Updated"] <- NA

pd <- position_dodge(1) 
ggplot(dfm_r_ag, aes(x=n_trials, y=Mean, color = variable, fill = variable, group = variable)) + 
    geom_errorbar(aes(ymin=ci_L, ymax=ci_U), width=.2, position=pd) + geom_line(position=pd) + geom_point(position=pd) + coord_cartesian(ylim = c(0,1), xlim = c(0,80)) + my_opts +  geom_hline(yintercept=.9, color='black', linetype = 2) + geom_hline(yintercept=.8, color='black', linetype = 1) + geom_hline(yintercept=.7, color='black', linetype = 3) + geom_hline(yintercept=.6, color='red', linetype = 4)  + facet_wrap(~Age)

ggplot(dfm_r_ag, aes(x=n_trials, y=Mean, color = as.factor(Age), fill = as.factor(Age), group = as.factor(Age))) + 
    geom_errorbar(aes(ymin=ci_L, ymax=ci_U), width=.3, position=pd) + geom_line(position=pd) + geom_point(position=pd) + coord_cartesian(ylim = c(0,1), xlim = c(0,80)) + my_opts +  geom_hline(yintercept=.9, color='black', linetype = 2) + geom_hline(yintercept=.8, color='black', linetype = 1) + geom_hline(yintercept=.7, color='black', linetype = 3) + geom_hline(yintercept=.6, color='red', linetype = 4)  + facet_wrap(~variable)

dfm_r_ag$Mean_fix <- ifelse(dfm_r_ag$Mean <0, 0, dfm_r_ag$Mean)
dfm_r_ag$ci_L_fix <- ifelse(dfm_r_ag$ci_L <0, 0, dfm_r_ag$ci_L)
dfm_r_ag$ci_U_fix <- ifelse(dfm_r_ag$ci_U <0, 0, dfm_r_ag$ci_U)
dfm_r_ag$variable_fix <- recode_factor(dfm_r_ag$variable, !!!list('Correct_ERN'='Correct (CRN)', 'Correct_Pe'='Correct (Pe)', 
                                                                  'Error_ERN'='Error (ERN)', 'Error_Pe'='Error (Pe)')) # Recoding for plot

dfm_r_ag$Age <- as.factor(dfm_r_ag$Age)

(p.erp.rel <- ggplot(dfm_r_ag, aes(x=n_trials, y=Mean_fix, color = Age, fill = Age, group = Age)) + 
    geom_errorbar(aes(ymin=ci_L_fix, ymax=ci_U_fix), width=.3, position=pd) + geom_line(position=pd) + geom_point(position=pd) + coord_cartesian(ylim = c(0,1), xlim = c(0,32)) + my_opts +  geom_hline(yintercept=.9, color='black', linetype = 2) + geom_hline(yintercept=.8, color='black', linetype = 1) + geom_hline(yintercept=.7, color='black', linetype = 3) + geom_hline(yintercept=.6, color='red', linetype = 4) + labs(x = "Number of Trials", y = "Split-Half Reliability of ERPs") +  scale_x_continuous(limits = c(0, 34), breaks = seq(0, 32, by = 4)) + theme(legend.position="none") + facet_wrap(~variable_fix) )

# Getting the number of trials needed for a given cutoff
dfm_r_ag_cutoff_mean <- dfm_r_ag %>% 
  group_by(variable, Age) %>% # Grouping by vars of interest
  filter(Mean >= .6) %>%
  filter(n_trials == min(n_trials)) %>%
  ungroup()

dfm_r_ag_cutoff_mean %>% select(n_trials, variable_fix, Age) %>% spread(variable_fix, n_trials) 

dfm_r_ag_cutoff_ci <- dfm_r_ag %>% 
  group_by(variable, Age) %>% # Grouping by vars of interest
  filter(ci_U >= .6) %>%
  filter(n_trials == min(n_trials)) %>%
  ungroup()

dfm_r_ag_cutoff_ci %>% select(n_trials, variable_fix, Age) %>% spread(variable_fix, n_trials) 

# ###### Checking the power estimates to see how many trials for a significant effect
# # Making sure we do not have less than six participants
# df_t$t_val_Deviant[df_t$df_Deviant < (cutoff-2)] <- NA # It is dfs so I need to subtract two 
# df_t$t_val_Deviant_late[df_t$df_Deviant_late < (cutoff-2)] <- NA # It is dfs so I need to subtract two 
# df_t$t_val_Novel[df_t$df_Novel < (cutoff-2)] <- NA # It is dfs so I need to subtract two 
# 
# dfm_t <- melt(df_t[,c("t_val_Deviant", "t_val_Deviant_late", "t_val_Novel", "n_trials", "Age")], id=c("n_trials", "Age"))
# dfm_t$variable <- recode_factor(dfm_t$variable, !!!list('t_val_Deviant'='Deviant', 't_val_Deviant_late'='Deviant Late', 't_val_Novel'='Novel')) # Recoding for plot
# 
# # Plotting
# ggplot(dfm_t, aes(n_trials, abs(value), color = variable, fill = variable, group = variable)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line",  aes(group = variable)) + my_opts+ labs(x = "Number of Trials", y = "t-value") + geom_hline(yintercept=1.96, color='black', linetype = 2) + geom_hline(yintercept=3.291, color='black', linetype = 1)  + theme(legend.position="bottom", legend.title = element_blank()) + facet_wrap(~Age)
# 
# # Power analyses - how many random subsamples cross the limit? 
# df_t_per <- dfm_t %>% mutate(over_05 = ifelse(abs(value) > 1.96, 1, 0),
#                          over_001 = ifelse(abs(value) > 3.291, 1, 0)) %>%
#         group_by(n_trials, variable, Age) %>%
#         summarise(over_05 = sum(over_05)/length(over_05),
#                   over_001 = sum(over_001)/length(over_001))
# names(df_t_per)[names(df_t_per) == "variable"] <- "type"
# df_t_per <- melt(df_t_per, id.vars = c("n_trials", "type", "Age"))  
# df_t_per$variable <- recode_factor(df_t_per$variable, !!!list('over_05'='t=1.96 (p<.05)', 'over_001'='t=3.29 (p<.001)')) # Recoding for plot
# 
# ggplot(df_t_per, aes(n_trials, value, color = type, fill = type, group = type)) + stat_summary(fun.y = mean, geom = "point",  position=position_dodge(width=0.7)) + stat_summary(fun.y = mean, geom = "line",  aes(group = type)) + my_opts+ labs(x = "Number of Trials", y = "Percent of bootstrap samples") + facet_wrap(~variable) + geom_hline(yintercept=.9, color='black', linetype = 2) + geom_hline(yintercept=.8, color='black', linetype = 1) + geom_hline(yintercept=.95, color='black', linetype = 3) + facet_wrap(~Age)
# 
# 
# # It seems like this would be heavily impacted by sample size - it would be good to provide actual effect sizes. However, even Cohen's d is dependent on the sample size

#####
# Plotting the effect sizes
df_t$Cohend_ERN[df_t$df_ERN < (cutoff-2)] <- NA # It is dfs so I need to subtract two 
df_t$Cohend_Pe[df_t$df_Pe < (cutoff-2)] <- NA # It is dfs so I need to subtract two 

dfm_Cohend <- melt(df_t[,c("Cohend_ERN", "Cohend_Pe", "n_trials", "Age")], id=c("n_trials", "Age"))
dfm_Cohend$variable <- recode_factor(dfm_Cohend$variable, !!!list('Cohend_ERN'='ERN', 'Cohend_Pe'='Pe')) # Recoding for plot

# Plotting
ggplot(dfm_Cohend, aes(n_trials, abs(value), color = variable, fill = variable, group = variable)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line",  aes(group = variable)) + my_opts+ labs(x = "Number of Trials", y = "Effect Size (Cohen's d)") + theme(legend.position="bottom", legend.title = element_blank()) + facet_wrap(~Age)

dfm_Cohend$Age <- as.factor(dfm_Cohend$Age)
ggplot(dfm_Cohend, aes(n_trials, abs(value), color = Age, fill = Age, group = Age)) + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line",  aes(group = Age)) + my_opts+ labs(x = "Number of Trials", y = "Effect Size (Cohen's d)") + theme(legend.position="bottom", legend.title = element_blank()) + facet_wrap(~variable) + ylim(0, 1.7)

# Plots with confidence intervals
dfm_Cohend_ag <- dfm_Cohend %>% 
  group_by(n_trials, variable, Age) %>% # Grouping by vars of interest
  dplyr::summarise(Mean = mean(value, na.rm = T), Median = median(value), ci_L = quantile(value, .025, na.rm = T), ci_U = quantile(value, .975, na.rm = T))

dfm_Cohend_ag$Age <- as.factor(dfm_Cohend_ag$Age)

ggplot(dfm_Cohend_ag, aes(x=n_trials, y=Mean, color = Age, fill = Age, group = Age)) + 
    geom_errorbar(aes(ymin=ci_L, ymax=ci_U), width=.2, position=pd) + geom_line(position=pd) + geom_point(position=pd) + coord_cartesian(xlim = c(0,32)) + my_opts + labs(x = "Number of Trials", y = "Effect Size (Cohen's d)") + theme(legend.position="bottom", legend.title = element_blank()) + facet_wrap(~variable)

dfm_Cohend_ag$Age <- as.numeric(gsub("Y", "", dfm_Cohend_ag$Age))
dfm_Cohend_ag$Age <- as.factor(dfm_Cohend_ag$Age)

# Plotting
(p.erp.es <- ggplot(dfm_Cohend_ag, aes(n_trials, Mean, color = Age, fill = Age, group = Age)) + geom_line(position=pd) + geom_point(position=pd) +  geom_errorbar(aes(ymin=ci_L, ymax=ci_U), width=.2, position=pd) + geom_line(position=pd) + my_opts+ labs(x = "Number of Trials", y = "Effect Size of ERP (Cohen d)") +  scale_x_continuous(limits = c(0, 34), breaks = seq(0, 32, by = 4)) + theme(legend.position="none") + facet_wrap(~variable) +
     geom_hline(yintercept=.8, color='black', linetype = 2) + geom_hline(yintercept=.5, color='black', linetype = 1) + geom_hline(yintercept=.2, color='black', linetype = 3) +
    geom_hline(yintercept=-.8, color='black', linetype = 2) + geom_hline(yintercept=-.5, color='black', linetype = 1) + geom_hline(yintercept=-.2, color='black', linetype = 3) )


```

